Viscous Transport in Eroding Porous Media

### 2019-10-03 docs/transport/erosion_transport.tex
small change

Following our previous work~\cite{qua-moo2018}, we begin by analyzing
the effect of erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  The general trend of the area fraction
resembles our previous work~\cite{qua-moo2018} (Figure 10(a)), but with
a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as a exponential law.  The line of
best fit is
\begin{align} 
  \log_{10} U \approx 3.43 \frac{t}{t_f} + c, 
\end{align}
$t_f$ is the time required for all the bodies to erode, and $c$ is a
fitting parameter. By construction $U = 1$ when $t=t_f$, so $c=-3.43$,
and the flow rate is approximated as
\begin{align}
  U \approx \exp\left(7.90 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20dense}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20
nearly touching eroding bodies versus normalized time. (b) The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as a power law (dashed line) towards the flow rate
$U=1$ that occurs once all the bodies have eroded.}
\end{figure}

In Figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to outlet where the velocity and
vorticity, and therefore erosion rate, are much larger than other
regions.  These channels can be further visualized by considering the
streamlines.  In Figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from Figure~\ref{fig:Eroding20vort} and plot 200 streamlines that are initialized at $x=-1$ and equally
spaced in $y \in (-1,1)$.  The streamlines are shown at five different
times, and the final plot is a zoom in of the lower right quadrant of
the fifth time step.  Since we use a high-order quadrature rule and time
stepping method, we resolve streamlines that come very close to the
eroded bodies.  There are three clear regions where the streamlines are
most concentrated, corresponding to the regions of highest velocity.
Two of these regions are located between the bodies and the solid walls
at $y=\pm 1$, and the third cuts through the porous region with the
upper part of the channel formed by bodies 1, 4, 6, and 8.  Since the
flow is fastest in these regions, the shearing is largest, and this
causes the channels to continue to open fastest as observed in
Figure~\ref{fig:Eroding20vort}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in $y \in (-1,1)$. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot.  Since we use the Barycentric
quadrature rule and the fourth-order Runge-Kutta method, some
streamlines are very close to the grains without penetrating them.}
\end{center}
\end{figure}

>>>>

Following our previous work~\cite{qua-moo2018}, we begin by analyzing
the effect of erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  The general trend of the area fraction
resembles our previous work~\cite{qua-moo2018} (Figure 10(a)), but with
a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as a exponential law.  The line of
best fit is
\begin{align} 
  \log_{10} U \approx 4.32 \frac{t}{t_f} + c, 
\end{align}
$t_f$ is the time required for all the bodies to erode, and $c$ is a
fitting parameter. By construction $U = 1$ when $t=t_f$, so $c=-4,32$,
and the flow rate is approximated as
\begin{align}
  U \approx \exp\left(9.94 \left(\frac{t - t_f}{t_f} \right)\right),
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

\begin{figure}[H]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/porosity20dense}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.7\linewidth]{./figs/flow_rate20densen}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding20flowrate}(a) The area fraction of 20
nearly touching eroding bodies versus normalized time. (b) The flow
rate, $U$, for a fixed pressure drop across the channel versus
normalized time.  The flow rate is initially very small, but it
eventually increases as a power law (dashed line) towards the flow rate
$U=1$ that occurs once all the bodies have eroded.}
\end{figure}

In Figure~\ref{fig:Eroding20vort}, we observe that erosion creates a
network of channels from the inlet to outlet where the velocity and
vorticity, and therefore erosion rate, are much larger than other
regions.  These channels can be further visualized by considering the
streamlines.  In Figure~\ref{fig:Eroding20tracer}, we freeze the
geometry at the second snapshot from Figure~\ref{fig:Eroding20vort} and plot 200 streamlines that are initialized at $x=-1$ and equally
spaced in $y \in (-1,1)$.  The streamlines are shown at five different
times, and the final plot is a zoom in of the lower right quadrant of
the fifth time step, but with additional streamlines.  Since we use a
high-order quadrature rule and time stepping method, we resolve
streamlines that come very close to the eroded bodies.  There are three
clear regions where the streamlines are most concentrated, corresponding
to the regions of highest velocity.  Two of these regions are located
between the bodies and the solid walls at $y=\pm 1$, and the third cuts
through the porous region with the upper part of the channel formed by
bodies 1, 4, 6, and 8.  Since the flow is fastest in these regions, the
shearing is largest, and this causes the channels to continue to open
fastest as observed in Figure~\ref{fig:Eroding20vort}.

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b30n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b90n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b150n}\\
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b210n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n}
\includegraphics[width = 0.32 \textwidth]{./figs/tracer_20b270n_zoom}
\caption{\label{fig:Eroding20tracer} 200 streamlines in the second
geometry from Figure~\ref{fig:Eroding20vort}. The streamlines are
initialized at $x=-1$ and are equally spaced in $y \in (-1,1)$. The
first five snapshots are evenly spaced in time.  The bottom right frame
is a zoom in of the fifth snapshot, but with additional streamlines.
Since we use the Barycentric quadrature rule and the fourth-order
Runge-Kutta method, some streamlines are very close to the grains
without penetrating them.}
\end{center}
\end{figure}
